---
title: "Lab 1"
author: "Shayan Kaveh"
date: "1/5/2022"
output: html_document
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
# can delete include=TRUE because it's the default
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
### install.packages('tmap')
### update.packages(ask = FALSE)
```

### Read in the data

cmd-option-i shortcut for creating a code shunk

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"),
                     show_col_types = FALSE)
```

### Part 1: wrangling and ggplot review

**Example 1:** Find counts of observation by `legal_status` & wrangle a bit.

```{r}
### method 1: group_by %>% summarize()
sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_count = n())

### method 2: different way plus a few new functions
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
# relocate() helps you reorder columns
  slice_max(tree_count, n = 5) %>% 
  arrange(-tree_count) # add - to flip the order; or add arrange(des(tree_count))
```

Make a graph of the top 5 from above:

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col(fill = "darkgreen") +
  labs(x = "legal status", y = "tree count") +
  coord_flip() +
  theme_minimal()
```

**Example 2:** Only going to keep observations where legal status is "Permitted Site" and caretaker is "MTA", and store as `permitted_data_df`

shift-cmmd-c to turn text into comment!!! 
```{r}
# sf_trees$legal_status %>% unique()
permitted_data_df <- sf_trees %>%
  filter(legal_status == "Permitted Site", caretaker == "MTA")
```

**Example 3:** Only keep Blackwood Acacia Trees, and then only keep colums `legal_status`, `data`, `latitude`, `longitude` and store as `blackwood_acacia_df`

```{r}
blackwood_acacia_df <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, lat = latitude, lon = longitude) 

### Make a little graph of locations
ggplot(data = blackwood_acacia_df, aes(x = lon, y = lat)) +
  geom_point(color = "darkgreen")
```

**Example 4:** use `tidyr::separate()`

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c('spp_scientific', 'spp_common'), sep = ' :: ')
```

**Example 5:** use `tidyr::unite()`

```{r}
ex_5 <- sf_trees %>% 
  unite("id_status", tree_id, legal_status, sep = "_COOL_")
```



